Structural replication of nanoporous media using procedural noise

### 2018-05-29 ms.tex
More fixes

\subsection{Multiple octaves: creating patterns}
\label{sec:octaves}
The different octaves from procedural noise functions can be combined
to produce a wide range of nature-like patterns. A generic
$N$-dimensional procedural function $\Phi(\vec x): \mathbb R^N \to
\mathbb R$ could be expressed as
\begin{equation}
\label{eq:procedural}
 \Phi(\vec x) = \Theta \Big(\sum_k^\Omega \kappa (k) P(  f_s (\vec x,k) )) \Big),
 \end{equation}
where $f_s(\vec x,k)$ describes the frequency scale modifier (e.g. $f_s(\vec x,k) =
k\vec x$), $\Omega$ is the amount of octaves (modes), $\kappa(k)$ is the octave amplitude (e.g. $\kappa(k) = 1/k$)
and $\Theta(x)$ an overall modifier (e.g. $\Theta(x) = x$ or
$\Theta(x) = \frac{1}{x}$). In a sense, $\kappa(k)$ is similar to Fourier coefficients. 
In its simplest case, summing directly over the octaves with amplitude $1/k$ results in 
\begin{equation}
\label{eq:perlinlinear}
 \Phi(\vec x) = \sum_k^\Omega \frac{1}{k} P( 2k\vec x).
\end{equation}
In one dimension, we show equation \ref{eq:perlinlinear} with 1, 3 and 30 octaves
in figure \ref{fig:1dperlin}, while the 2D procedural noise pattern is depicted in the lower
part of figure \ref{fig:noise}. Finally, since simplex noise both scales better
with higher dimensions and produces less asymmetric effects, we choose to use
simplex noise over regular Perlin noise as the main noise function throughout this paper.

@@

\begin{figure*}
\includegraphics[width=.95\textwidth]{model_examples.png}
\caption{Examples of 3D simplex noise structures in nanoporous media. Left image:
low persistence (large scale dominates), middle image: medium persistence (small
scales starts dominating) and right image: high threshold (more atoms are discarded). }
\label{fig:model_example}
\end{figure*}
We construct a generic sum over procedural noise octaves with the expression: 
\begin{equation}
  M_1(\bar \theta) = M_1(\vec x, \psi, \Omega, \nu) = \sum_{\omega=0}^{\Omega} \psi^{\omega -1}   N\big((\vec x + \vec\sigma \cdot \omega)\nu \cdot 2^{\omega-1} \big),
\label{eq:noisemodel1}
\end{equation}
where $\Omega$ is the number of octaves, $\psi$ the persistence (scale falloff),
$N$ a generic $N$-dimensional noise function, $\vec \sigma$ a predefined random
vector that removes anisotropic artefacts and $\nu$ the frequency modifier representing
the initial largest scales in the model. For simplicity, $\bar \theta$ here denotes
the set of model parameters. For each mode, the frequency is thus multiplied, while
the overall amplitude is damped by a given persistence. In the end, we simulate a
large set of bulk particles without any walls, and use the noise function to remove
particles that are not within a procedural void given by a threshold $\tau$. Three
examples of this model are depicted in figure \ref{fig:model_example}, where we have
used $\Omega=6$ octaves and the scale is $\nu=0.01$ and cut-off $\tau=0$. The left
image depicts low persistence ($\psi = 0.0$), where only the first octave dominates
(corresponding to large scales). The middle picture shows the same parameters with
a higher persistence $(\psi = 0.5)$, where smaller scales become more dominating.
In the rightmost picture, the effect of the cut-off threshold $\tau = -0.3$ is shown to emphasize threshold cut-off.

@@

Ideally, the model measure should pick up all geometrical properties that are
relevant for the physical study of interest. 
For instance, transport in porous media heavily depends on the pore size
distribution and porosity, whereas chemical reactons are sensitive to surface area \cite{coussy2011mechanics}.
In this paper, we choose to focus on the radial distribution function (RDF)
$g(r)$ as defined and implemented in LAMMPS \cite{plimpton1995fast}.
$g(r)$ describes how density varies as a function of distance from a 
reference particle, picking up both the intramolecular and intermolecular structure. 
We have chosen to ignore the short range contributions of $g(r)$ since
the pore geometry does not depend on the intra-molecular structure of
SiO$_2$, but only the density variations at larger distances. 
Hereafter, $g_d(r)$ denotes the RDF calculated from simulated or experimental data,
while $g_{M(\bar \theta)}(r)$ represents the RDF based on the procedural noise
model $M(\bar \theta)$, where $\bar \theta$ is the set of model parameters.

The method used for calculating porosity of a media is defined as in \cite{gelb1998characterization},
and we obtain the surface area of a porous media by using m arching cubes triangulation to sum the
surface area of the generated triangles.
As discussed in \cite{gelb1998characterization}, the definition of surface area is ill-defined in
atomic simulations, because a clear definition of pore space is needed. 
In the marching cubes triangulation, we create a voxelized grid, a scalar field with values being
the distance from the voxel centre to the nearest atom. 
The isovalue threshold defining the surface is chosen to be \SI{2}{\angstrom}, large enough to
contribute to a dense pore matrix, but small enough to have most of the available pore space defined as such. 
The important note here is that the same definition of both surface and porosity is being used
in all systems that are compared, physical simulations, procedural methods or data.

>>>>

\subsection{Multiple octaves: creating patterns}
\label{sec:octaves}
The different octaves from procedural noise functions can be combined
to produce a wide range of nature-like patterns. A generic
$N$-dimensional procedural function $\Phi(\vec r): \mathbb R^N \to
\mathbb R$ could be expressed as
\begin{equation}
\label{eq:procedural}
 \Phi(\vec r) = \Theta \Big(\sum_k^\Omega \kappa (k) P(  f_s (\vec r,k) )) \Big),
 \end{equation}
where $f_s(\vec r,k)$ describes the frequency scale modifier (e.g. $f_s(\vec r,k) =
k\vec r$), $\Omega$ is the amount of octaves (modes), $\kappa(k)$ is the octave amplitude (e.g. $\kappa(k) = 1/k$)
and $\Theta(x)$ an overall modifier (e.g. $\Theta(x) = x$ or
$\Theta(x) = \frac{1}{x}$). In a sense, $\kappa(k)$ is similar to Fourier coefficients. 
In its simplest case, summing directly over the octaves with amplitude $1/k$ results in 
\begin{equation}
\label{eq:perlinlinear}
 \Phi(\vec r) = \sum_k^\Omega \frac{1}{k} P( 2k\vec r).
\end{equation}
In one dimension, we show equation \ref{eq:perlinlinear} with 1, 3 and 30 octaves
in figure \ref{fig:1dperlin}, while the 2D procedural noise pattern is depicted in the lower
part of figure \ref{fig:noise}. Finally, since simplex noise both scales better
with higher dimensions and produces less asymmetric effects, we choose to use
simplex noise over regular Perlin noise as the main noise function throughout this paper.

@@

\begin{figure*}
\includegraphics[width=.95\textwidth]{model_examples.png}
\caption{Examples of 3D simplex noise structures in nanoporous media. Left image:
low persistence (large scale dominates), middle image: medium persistence (small
scales starts dominating) and right image: high threshold (more atoms are discarded). }
\label{fig:model_example}
\end{figure*}
We construct a generic sum over procedural noise octaves with the expression: 
\begin{equation}
  M_1(\vec r |\bar \theta) = M_1(\vec r | \psi, \Omega, \nu) = \sum_{\omega=0}^{\Omega} \psi^{\omega -1}   N\big((\vec r + \vec\sigma \cdot \omega)\nu \cdot 2^{\omega-1} \big),
\label{eq:noisemodel1}
\end{equation}
where $\Omega$ is the number of octaves, $\psi$ the persistence (scale falloff),
$N$ a generic $N$-dimensional noise function, $\vec \sigma$ a predefined random
vector that removes anisotropic artefacts and $\nu$ the frequency modifier representing
the initial largest scales in the model. For simplicity, $\bar \theta$ here denotes
the set of model parameters. For each mode, the frequency is thus multiplied, while
the overall amplitude is damped by a given persistence.

We then get, for a given position $\vec r$, a scalar value from the noise function.
Assume now that we have a system in molecular dynamics where all atomic positions are given.
A porous material is then generated by introducing a threshold $\tau$ which is used to decide whether or not an atom should be removed from the system.
A boolean function $\text{Delete}(\vec r)$ is defined as

\begin{align}
  \text{Delete}(\vec r) = 
  \begin{cases} 
    \text{True} & \text{if} \quad M_1(\vec r |\bar \theta) < \tau \\
    \text{False} & \text{if} \quad M_1(\vec r |\bar \theta) \geq \tau,
  \end{cases}
\end{align}
where we simply evaluate this function for every atom in the system.
The threshold $\tau$ is now added as a parameter to the model.
Three examples of this model are depicted in figure \ref{fig:model_example}, where we have
used $\Omega=6$ octaves and scale $\nu=0.01$ with threshold $\tau=0$. The left
image depicts low persistence ($\psi = 0.0$), where only the first octave dominates
(corresponding to large scales). The middle picture shows the same parameters with
a higher persistence $(\psi = 0.5)$, where smaller scales become more dominating.
In the rightmost picture, the effect of the threshold $\tau = -0.3$ is shown to emphasize threshold cut-off.

@@

Ideally, the model measure should pick up all geometrical properties that are
relevant for the physical study of interest. 
For instance, transport in porous media heavily depends on the pore size distribution and porosity\cite{kozeny1927uber, carman1937fluid},
whereas chemical reactons are sensitive to surface area\cite{coussy2011mechanics}.
In this paper, we choose to focus on the radial distribution function (RDF)
$g(r)$ as defined and implemented in LAMMPS\cite{plimpton1995fast}.
$g(r)$ describes how density varies as a function of distance from a 
reference particle, picking up both the intramolecular and intermolecular structure. 
We have chosen to ignore the short range contributions of $g(r)$ since
the pore geometry does not depend on the intra-molecular structure of
SiO$_2$, but only the density variations at larger distances. 
Hereafter, $g_d(r)$ denotes the RDF calculated from simulated or experimental data,
while $g_{M(\bar \theta)}(r)$ represents the RDF based on the procedural noise
model $M(\bar \theta)$, where $\bar \theta$ is the set of model parameters.

The method used for calculating porosity of a porous media is the same as in \cite{gelb1998characterization}.
Surface area is obtained by creating a voxelized grid of the system with values being the distance from each voxel center to the nearest atom.
This is now a scalar field and we can apply marching cubes to triangulate an isosurface of the distance field.
This gives a trianglated representation of the porous media defined by a set of atomic positions.
As discussed in \cite{gelb1998characterization}, the definition of surface area is ill-defined in
atomistic simulations, because a clear definition of pore space is difficult.
We have chosen to define the pore space as the set of points being more than \SI{2}{\angstrom} to any atom.
This is large enough to contribute to a dense pore matrix, but small enough to have most of the available pore space defined as such. 
The important note here is that the same definition of both surface and porosity is being used in all systems that are compared, physical simulations, procedural methods or data.
